Understanding Semiparametric Constraints

### 2024-01-23 Notes/semiconstraints.qmd
Automatically updated on Tue Jan 23 10:28:25 AM CST 2024

This is a similar framework to understanding the class of estimators 
that we discussed in \verb|optest_update.pdf|, where we write the estimator 
as a weighted average of different expectation with functional coefficients.

$$\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].$$

@@

$$\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}$$

where $\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1,
Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
\gamma_{kt}$ in @tbl-reftab.

@@

>>>>

This is a similar framework to understanding the class of estimators 
that we discussed in \verb|optest_update.pdf|, where we write the estimator 
as a weighted average of different expectations with functional coefficients.

$$\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta)E[g(Z)
\mid X] + \beta_1(\delta)E[g(Z) \mid X, Y_1] + \beta_2(\delta) E[g(Z)
\mid X, Y_2].$$

In this case, we assume that $E[\beta_j(\delta)] = 0$ for $j = 0, 1, 2$.

@@

$$\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}$$ {#eq-linest}

where $\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1,
Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
\gamma_{kt}$ in @tbl-reftab. In general the estimator has the following form,

\begin{align*}
\hat \theta = n^{-1} \sum_{i = 1}^n 
  &\left\{c_{11}\frac{\delta_{00}}{\pi_{00}} E[g \mid X]+ \right.\\
  &c_{21}\frac{\delta_{10}}{\pi_{10}} E[g \mid X]+ 
  c_{22}\frac{\delta_{10}}{\pi_{10}} E[g \mid X, Y_1]+ \\
  &c_{31}\frac{\delta_{01}}{\pi_{01}} E[g \mid X]+ 
  c_{33}\frac{\delta_{01}}{\pi_{01}} E[g \mid X, Y_2]+ \\
  &c_{41}\frac{\delta_{11}}{\pi_{11}} E[g \mid X]+ 
  c_{42}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1]+
  c_{43}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_2]+
  \left.c_{44}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1, Y_2]\right\}
\end{align*}

which we have condensed to  @eq-linest.

@@

We can justify the constraints for the Outcome Robust model by rewriting the 
estimator into the following form:

\begin{align*}
E_\delta[\hat \theta] &= n^{-1} \sum_{i = 1}^n E_\delta\left[
E[g \mid X] \left(\frac{\delta_{00}}{\pi_{00}} c_{11} +
\frac{\delta_{10}}{\pi_{10}}c_{21} \frac{\delta_{01}}{\pi_{01}}c_{31} +
\frac{\delta_{11}}{\pi_{11}}c_{41}\right)\right. + \\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_1] \left(\frac{\delta_{10}}{\pi_{10}}c_{22} + 
\frac{\delta_{11}}{\pi_{11}}c_{42}\right) +\\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_2] \left(\frac{\delta_{01}}{\pi_{01}}c_{33} + 
\frac{\delta_{11}}{\pi_{11}}c_{43}\right) +\\ 
&\qquad\qquad\qquad\qquad \left.g \frac{\delta_{11}}{\pi_{11}}c_{44}\right] \\
&= n^{-1} \sum_{i = 1}^n \left\{E[g \mid X](c_{11} + c_{21} + c_{31} + c_{41}) +
E[g \mid X, Y_1] (c_{22} + c_{42}) +
E[g \mid X, Y_2] (c_{33} + c_{43}) + g c_{44}\right\}
\end{align*}

Thus, if the Outcome Robust constraints are satisfied, 
$E_\delta[\hat \theta] = g$ 
when we take the expectation with respect to the response variable $\delta$.
These constraints make sense because if they are satisfied $E[\hat \theta]$ 
does not depend on the correct specification of the models
for $E[g \mid G_r(X, Y_1, Y_2)]$.

Likewise, we can justify the constraints for the Response Robust model 
by rewriting @eq-linest into the form

\begin{align*}
E_Z [\hat \theta] &= n^{-1} \sum_{i = 1}^n E_Z\left[
\frac{\delta_{00}}{\pi_{00}}c_{11} E[g \mid X]\right. + \\
&\qquad\qquad\qquad\qquad\frac{\delta_{10}}{\pi_{10}}\left(c_{21}E[g \mid X] + 
  c_{22}E[g \mid X, Y_1]\right) +\\
&\qquad\qquad\qquad\qquad\frac{\delta_{01}}{\pi_{01}}\left(c_{31}E[g \mid X] + 
  c_{33}E[g \mid X, Y_2]\right) +\\
&\qquad\qquad\qquad\qquad\left.\frac{\delta_{11}}{\pi_{11}}\left(c_{41}E[g \mid X] + 
  c_{42}E[g \mid X, Y_1] + c_{43}E[g \mid X, Y_2] + c_{44}g\right)\right] \\
& n^{-1} \sum_{i = 1}^n \left(\frac{\delta_{00}}{\pi_{00}}c_{11} + 
\frac{\delta_{10}}{\pi_{10}}(c_{21} + c_{22}) + 
\frac{\delta_{01}}{\pi_{01}}(c_{31} + c_{33}) + 
\frac{\delta_{11}}{\pi_{11}}(c_{41} + c_{42} + c_{43} + c_{44})\right)g
\end{align*}

Thus, if the Response Robust constraints are satisfied, $E_X[\hat \theta] = g$
when we take the expectation with respect to the outcome variable 
$Z = (X, Y_1, Y_2)'$. These constraints make sense because when they are 
satisfied $E[\hat \theta]$ does not depend on the correct response 
probabilities $\pi = E[\delta]$.

### 2024-01-23 Notes/semiconstraints.tex
Automatically updated on Tue Jan 23 10:28:25 AM CST 2024

\title{Understanding Semiparametric Constraints}
\author{Caleb Leedy}
\date{19 January 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

This is a similar framework to understanding the class of estimators
that we discussed in \verb|optest_update.pdf|, where we write the
estimator as a weighted average of different expectation with functional
coefficients.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].\]

@@

\[\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}\]

where
\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
and \(ij\) corresponds to the segment \(A_{ij}\) associated with
\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}.

@@

@@

\hypertarget{tbl-fstat}{}
\begin{longtable}[]{@{}llrr@{}}
\caption{\label{tbl-fstat}This table computes the F-statistic and
P-value for an F-test comparing the \texttt{Full\ Model} with a nested
model in \texttt{Reduced\ Model}. The computation of the F-statistic is
found in the \texttt{get\_f\_statistic} function.}\tabularnewline
\toprule\noalign{}
Full Model & Reduced Model & Fstat & P-Value \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Full Model & Reduced Model & Fstat & P-Value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & Parametric 2 & 0.00 & 1 \\
Parametric 1 & Outcome Robust & 1190.30 & 0 \\
Parametric 1 & Response Robust & 0.00 & 1 \\
Parametric 1 & Double Robust & 510.13 & 0 \\
Parametric 2 & Outcome Robust & 1787.26 & 0 \\
Parametric 2 & Response Robust & 0.00 & 1 \\
Parametric 2 & Double Robust & 595.75 & 0 \\
Outcome Robust & Double Robust & 0.00 & 1 \\
Response Robust & Double Robust & 895.43 & 0 \\
\end{longtable}

>>>>

\title{Understanding Semiparametric Constraints}
\author{Caleb Leedy}
\date{22 January 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi

@@

This is a similar framework to understanding the class of estimators
that we discussed in \verb|optest_update.pdf|, where we write the
estimator as a weighted average of different expectations with
functional coefficients.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta)E[g(Z)
\mid X] + \beta_1(\delta)E[g(Z) \mid X, Y_1] + \beta_2(\delta) E[g(Z)
\mid X, Y_2].\]

In this case, we assume that \(E[\beta_j(\delta)] = 0\) for
\(j = 0, 1, 2\).

@@

\begin{equation}\protect\hypertarget{eq-linest}{}{\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}}\label{eq-linest}\end{equation}

where
\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
and \(ij\) corresponds to the segment \(A_{ij}\) associated with
\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}. In general the estimator
has the following form,

\begin{align*}
\hat \theta = n^{-1} \sum_{i = 1}^n 
  &\left\{c_{11}\frac{\delta_{00}}{\pi_{00}} E[g \mid X]+ \right.\\
  &c_{21}\frac{\delta_{10}}{\pi_{10}} E[g \mid X]+ 
  c_{22}\frac{\delta_{10}}{\pi_{10}} E[g \mid X, Y_1]+ \\
  &c_{31}\frac{\delta_{01}}{\pi_{01}} E[g \mid X]+ 
  c_{33}\frac{\delta_{01}}{\pi_{01}} E[g \mid X, Y_2]+ \\
  &c_{41}\frac{\delta_{11}}{\pi_{11}} E[g \mid X]+ 
  c_{42}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1]+
  c_{43}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_2]+
  \left.c_{44}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1, Y_2]\right\}
\end{align*}

which we have condensed to Equation~\ref{eq-linest}.

@@

We can justify the constraints for the Outcome Robust model by rewriting
the estimator into the following form:

\begin{align*}
E_\delta[\hat \theta] &= n^{-1} \sum_{i = 1}^n E_\delta\left[
E[g \mid X] \left(\frac{\delta_{00}}{\pi_{00}} c_{11} +
\frac{\delta_{10}}{\pi_{10}}c_{21} \frac{\delta_{01}}{\pi_{01}}c_{31} +
\frac{\delta_{11}}{\pi_{11}}c_{41}\right)\right. + \\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_1] \left(\frac{\delta_{10}}{\pi_{10}}c_{22} + 
\frac{\delta_{11}}{\pi_{11}}c_{42}\right) +\\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_2] \left(\frac{\delta_{01}}{\pi_{01}}c_{33} + 
\frac{\delta_{11}}{\pi_{11}}c_{43}\right) +\\ 
&\qquad\qquad\qquad\qquad \left.g \frac{\delta_{11}}{\pi_{11}}c_{44}\right] \\
&= n^{-1} \sum_{i = 1}^n \left\{E[g \mid X](c_{11} + c_{21} + c_{31} + c_{41}) +
E[g \mid X, Y_1] (c_{22} + c_{42}) +
E[g \mid X, Y_2] (c_{33} + c_{43}) + g c_{44}\right\}
\end{align*}

Thus, if the Outcome Robust constraints are satisfied,
\(E_\delta[\hat \theta] = g\) when we take the expectation with respect
to the response variable \(\delta\). These constraints make sense
because if they are satisfied \(E[\hat \theta]\) does not depend on the
correct specification of the models for \(E[g \mid G_r(X, Y_1, Y_2)]\).

Likewise, we can justify the constraints for the Response Robust model
by rewriting Equation~\ref{eq-linest} into the form

\begin{align*}
E_Z [\hat \theta] &= n^{-1} \sum_{i = 1}^n E_Z\left[
\frac{\delta_{00}}{\pi_{00}}c_{11} E[g \mid X]\right. + \\
&\qquad\qquad\qquad\qquad\frac{\delta_{10}}{\pi_{10}}\left(c_{21}E[g \mid X] + 
  c_{22}E[g \mid X, Y_1]\right) +\\
&\qquad\qquad\qquad\qquad\frac{\delta_{01}}{\pi_{01}}\left(c_{31}E[g \mid X] + 
  c_{33}E[g \mid X, Y_2]\right) +\\
&\qquad\qquad\qquad\qquad\left.\frac{\delta_{11}}{\pi_{11}}\left(c_{41}E[g \mid X] + 
  c_{42}E[g \mid X, Y_1] + c_{43}E[g \mid X, Y_2] + c_{44}g\right)\right] \\
& n^{-1} \sum_{i = 1}^n \left(\frac{\delta_{00}}{\pi_{00}}c_{11} + 
\frac{\delta_{10}}{\pi_{10}}(c_{21} + c_{22}) + 
\frac{\delta_{01}}{\pi_{01}}(c_{31} + c_{33}) + 
\frac{\delta_{11}}{\pi_{11}}(c_{41} + c_{42} + c_{43} + c_{44})\right)g
\end{align*}

Thus, if the Response Robust constraints are satisfied,
\(E_X[\hat \theta] = g\) when we take the expectation with respect to
the outcome variable \(Z = (X, Y_1, Y_2)'\). These constraints make
sense because when they are satisfied \(E[\hat \theta]\) does not depend
on the correct response probabilities \(\pi = E[\delta]\).

@@

\hypertarget{tbl-fstat}{}
\begin{longtable}[]{@{}llrr@{}}
\caption{\label{tbl-fstat}This table computes the F-statistic and
P-value for an F-test comparing the \texttt{Full\ Model} with a nested
model in \texttt{Reduced\ Model}. The computation of the F-statistic is
found in the \texttt{get\_f\_statistic} function.}\tabularnewline
\toprule\noalign{}
Full Model & Reduced Model & Fstat & P-Value \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Full Model & Reduced Model & Fstat & P-Value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & Parametric 2 & 0.00 & 1 \\
Parametric 1 & Outcome Robust & 799.85 & 0 \\
Parametric 1 & Response Robust & 0.00 & 1 \\
Parametric 1 & Double Robust & 342.79 & 0 \\
Parametric 2 & Outcome Robust & 1200.99 & 0 \\
Parametric 2 & Response Robust & 0.00 & 1 \\
Parametric 2 & Double Robust & 400.33 & 0 \\
Outcome Robust & Double Robust & 0.00 & 1 \\
Response Robust & Double Robust & 601.70 & 0 \\
\end{longtable}
